---
title: "Cereal EP"
author: "Alex Bystrov, Rocio Rodriguez, Bodam Jerry"
date: "2025-02-18"
output: cereal_filtered_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}

#Initial setup

rm(list=ls())

if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

pacman::p_load(
  fixest, haven, modelsummary, tidyverse, readxl, dplyr, plm, stargazer, IDPmisc, lfe, here, sandwich, AER, 
  dynlm, forecast, scales, quantmod, urca, nlme, lmtest, pdynmc, dynpanel, rstudioapi, pgmm, xtable, estimatr, 
  zoo, skimr, lubridate, patchwork
)


# Get the directory path of the current R Markdown file
current_rmd <- rstudioapi::getSourceEditorContext()$path

# Set the working directory to the location of the R Markdown file
setwd(dirname(current_rmd))


cereal <- read.csv(file.path(getwd(), "products.csv"), sep = ",")
# cement$date <- lubridate::as_date(cement$date)

```

# 1. A first look at the data. 

```{r pressure, echo=FALSE}

print(head(cereal))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r pressure, echo=FALSE}
plot(pressure)
```
vtv
vtvtv
vtvtv

```{r pressure, echo=FALSE}
plot(pressure)
```